---
title: 'Dissertation: "Leveraging Evolutionary Changes for Software Process Quality"'
subtitle: "Replication Package for All Results, Figures, and Tables"
author:
- Sebastian Hönel
date: "Rendered on `r format(Sys.time(), '%B %d, %Y')`"
bibliography: ../inst/REFERENCES.bib
papersize: a4
fontsize: 10pt
geometry: left=2cm,right=2cm,top=2.5cm,bottom=2.5cm
urlcolor: blue
link-citations: yes
linkcolor: blue
header-includes:
- \PassOptionsToPackage{style=authoryear,maxcitenames=1,maxbibnames=99}{biblatex}
- \usepackage{amsmath}
- \usepackage{amssymb}
- \usepackage{bm}
- \usepackage{mathtools}
- \usepackage{xurl}
- \usepackage[nottoc]{tocbibind}
- \usepackage{subfig}

output:
  bookdown::pdf_document2:
    fig_caption: yes
    number_sections: yes
    toc: yes
    toc_depth: 6
    df_print: kable
    keep_tex: yes
    latex_engine: pdflatex
    citation_package: biblatex
  md_document:
    toc: yes
    toc_depth: 6
    df_print: kable
    variant: gfm
  html_document:
    number_sections: yes
    toc: yes
    toc_depth: 6
    toc_float: yes
    df_print: kable
  word_document:
    toc: yes
    toc_depth: 6

#abstract: "\\noindent Here goes the abstract."
---

\newcommand*\mean[1]{\overline{#1}}
\newcommand{\abs}[1]{\left\lvert\,#1\,\right\rvert}
\newcommand{\norm}[1]{\left\lVert\,#1\,\right\rVert}
\newcommand{\infdiv}[2]{#1\;\|\;#2}
\newcommand\argmax[1]{\underset{#1}{arg\,max}}
\newcommand\argmin[1]{\underset{#1}{arg\,min}}

\listoffigures
\listoftables

```{r echo=FALSE}
# Set this to false to show R code
#PAPER_MODE <- FALSE
```


```{r setup, include=FALSE}
library(knitr)
#knitr::opts_chunk$set(echo = !PAPER_MODE)
#knitr::opts_chunk$set(eval = interactive())
knitr::opts_chunk$set(dpi = 300)
opts_chunk$set(tidy = TRUE, tidy.opts = list(indent=2))

en_US.UTF8 <- "en_US.UTF-8"
Sys.setenv(LANG = en_US.UTF8)
Sys.setenv(LC_ALL = en_US.UTF8)
Sys.setenv(LC_COLLATE = en_US.UTF8)
Sys.setenv(LC_CTYPE = en_US.UTF8)
Sys.setenv(LC_MONETARY = en_US.UTF8)
Sys.setenv(LC_NUMERIC = en_US.UTF8)
Sys.setenv(LC_TIME = en_US.UTF8)
```

```{r echo=FALSE}
source(file = "./helpers.R")
```


# Preface {-}

This notebook represents the replication package for my dissertation [@honel2023_phdthesis].
Rendering this notebook will exactly reproduce all results, figures, and tables as used therein.
Some data was taken from a technical report [@honel2023_fdtr] and reused here.




# Source code density Expectation and Standard deviation


```{r cache=TRUE}
temp <- read.table("../data/360k_density.csv", header = TRUE)
quantile(temp$Density, probs = sort(c(.065,seq(0,.7,by=.1), .72504)))
```

```{r}
projects_sc <- readRDS(file = "../data/projects_sc.rds")$projects_sc
```

```{r}
scd_vals <- c()
scd_means <- c()
use_x <- seq(from = 0, to = 1, length.out = 1e5)

for (pname in names(projects_sc)) {
	scd_vals <- c(scd_vals, projects_sc[[pname]]$SCD(use_x))
	scd_means <- c(scd_means, cubature::cubintegrate(f = projects_sc[[pname]]$SCD, lower = 0, upper = 1)$integral)
}

`names<-`(x = c(mean(scd_vals), mean(scd_means), sd(scd_vals)), value = c("mean", "mean (integral)", "sd"))
```




# Automatic Calibration

We want to show some preliminary results for *Automatic Calibration* in the thesis.
Previously, we have applied the rank transform to develop a sort of scoring mechanism using joint cumulative probabilities [@ulan2018_qmio].
Most recently, the *rank transform* of this approach has been generalized by allowing to use any value as ideal value [@honel2022_qrs].


For the thesis, I want to show a concrete example based on project management.
For that, I will use some of the activities as were previously derived from an average manifestation of the Fire Drill using weighted mixtures [@honel2023_ecs].
Over the (normalized) project time, we now have a probability distribution for each activity.
There are six activities, three for source code and three for issue-tracking.
We could say that the weighted mixtures represent a "process model" for the Fire Drill, from a quantitative perspective.
However, I would not go this far at this point and call it a pattern, but rather a *typical accumulation*.
It would perhaps become a reliable process model if we were to add many more observations of projects that were affected by the Fire Drill in some degree.


## Loading the Weighted Mixtures

Nevertheless, for the sake of demonstrating automatic calibration, we will take the weighted mixtures and use an approximation of these and call it a process model here.
Figure \ref{fig:smooth-mixtures} shows the weighted mixture for each activity, as well as a simplified/smoothed version.
We will pick one or two of these activities for demonstrating automatic calibration.

```{r}
# Let's load the weighted mixtures:
weighted_mixtures <- readRDS("../data/weighted_mixtures.rds")
```

```{r}
rejection_sampling <- function(x_data, y_data, xlim = c(0, 1), num_x = 1e3) {
	tempf <- stats::approxfun(x = x_data, y = y_data, yleft = 0, yright = 0)#, rule = 2)
	use_x <- seq(from = xlim[1], to = xlim[2], length.out = num_x)
	use_y <- stats::runif(n = length(use_x), min = 0, max = max(y_data))
	use_x[use_y <= tempf(use_x)]
}


gaussian_kde_pdf <- function(data) {
	# Closed-form Gaussian KDE PDF:
	use_bw <- bw.SJ(x = data)
	kde_pdf <- Vectorize(function(x, h = use_bw) {
		1 / (length(data) * h) * sum(dnorm((x - data) / h))
	})
	
	# Cut the PDF so its integral [0,1] = 1
	the_int <- cubature::cubintegrate(f = kde_pdf, lower = 0, upper = 1, maxEval = 1e4)$integral
	return(function(x) {
		kde_pdf(x) / the_int
	})
}


make_smooth_dens <- function(org_densf, domain = c(0, 1), npoints = 15, span = 0.3) {
	use_x <- seq(from = domain[1], to = domain[2], length.out = npoints)
	y <- org_densf(use_x)
	temp <- stats::loess.smooth(x = use_x, y = y, span = span, family = "g", evaluation = 1e3)
	
	# Rejection-sampling from this curve:
	rej_samples <- rejection_sampling(
		x_data = ((temp$x - min(temp$x)) / (max(temp$x) - min(temp$x))), y_data = temp$y)
	
	gaussian_kde_pdf(rej_samples)
}
```

```{r smooth-mixtures, echo=FALSE, warning=FALSE, fig.width=7, fig.height=5, fig.cap="Weighted mixtures for all source code and issue-tracking activities."}
par(mfrow = c(3,3), oma = rep(0,4), mar = c(2,2,2,1))

use_x <- seq(from = 0, to = 1, length.out = 1e3)
for (idx in 1:length(names(weighted_mixtures))) {
	act <- names(weighted_mixtures)[idx]
	wm <- weighted_mixtures[[act]]
	tempf <- make_smooth_dens(org_densf = wm, span = .25)
	y_max <- max(c(wm(use_x), tempf(use_x)))
	
	curve2(func = wm, from = 0, to = 1, ylim = c(0, y_max), main = act, xlab = "")
	grid()
	curve2(func = tempf, from = 0, to = 1, col = "red", add = TRUE)
	if (idx == 2) {
		legend(x = 0, y = y_max, legend = c("Original", "Smoothed"), lty = 1, lwd = 2, col = c("black", "red"), bg = "white")
	}
}
```


## Prepare for Scoring

I think we should select the smoothed density for $\mathsf{REQ}$ from Figure \ref{fig:smooth-mixtures} as it makes for a good example.
Automatic calibration requires a (representative) sample of the space of all possible outcomes.
The larger the sample, the more accurate the scores will be.
So, what we require here, is a set of $\mathsf{REQ}$ activities **as they might occur in real-life**.
This is the point where one would either gather all observed instances or try to simulate them.
We will do the latter.
As we have no expectation as to how this activity might unfold in a random project, we will just generate random continuous probability distributions.


The next step is then to select and compute features.
The features should quantify the difference between the process model and any simulated process.
For simplicity, we will compute the **area**, **correlation**, and **symmetric Jensen--Shannon divergence** between process and -model, and we will do so on a few selected segments.
This will allow us then later to inspect individual scores for previously unseen processes.
Let's say we simulate ten thousand processes, then we will get a data frame with ten thousand rows and some columns (metric per $n$th-segment $1\ldots n$).


The second requirement for automatic calibration is an **ideal value** that allows us to transform any feature $F_i$ into a distance $D_i$.
We will do this using the absolute value function.
For an ideal value $d_i$, we obtain $D_i=\abs{\left(F_i-d_i\right)}$.
This distribution (and esp. its (C)CDF) is then used to transform and (non-)linearly scale a feature to become a score.
Since our features are distances (from the ideal), the scores based thereon will have the meaning *larger is better*.
The *utopian* ideal value is for a *correlation* is $\bm{1}$, and for the area between it is $\bm{0}$.
For the Jensen--Shannon divergence, it is $\bm{\infty}$, since we transform its values using the negative logarithm.
However, for practical reasons, we will use the largest value observed during automatic calibration, as it is unlikely to observe a smaller value in practice, given that we drew many samples.


## Calibration

Now for the actual calibration.
Let's generate a few thousand random processes and compute their segment-wise correlation with the smoothed $\mathsf{CP}$ activity.

First, we'll define some helper functions.

```{r}
get_smoothed_curve <- function(seed, npoints = 15, span = 0.35) {
	if (is.na(seed)) {
		stop("A seed is required.")
	}
	set.seed(seed = seed)
	
	x <- sort(c(0, 1, runif(n = npoints - 2)))
	y <- runif(n = length(x), min = 0, max = 1)
	temp <- loess.smooth(x = x, y = y, span = span, family = "g", evaluation = 1000)
	
	# Rejection-sampling from this curve:
	rej_samples <- rejection_sampling(
		x_data = ((temp$x - min(temp$x)) / (max(temp$x) - min(temp$x))), y_data = temp$y)
	
	gaussian_kde_pdf(rej_samples)
}


jsd_segment <- function(p, q, ext = c(0, 0.1), xtol = 1e-20) {
	cubature::cubintegrate(f = function(x) {
		p_ <- p(x)
		q_ <- q(x)
		m_ <- .5 * (p_ + q_)
		tryCatch({
			if (abs(p_) < xtol || abs(q_) < xtol || abs(m_) < xtol) 0 else .5 * p_ * log(p_ / m_) + .5 * q_ * log(q_ / m_)
		}, warning = function(w) {
			stop(paste0(p(x), " ; ", q(x), " ; ", m_))
		})
	}, lower = ext[1], upper = ext[2], maxEval = 1e3)$integral
}
```



```{r}
use_activity <- "REQ"
pm <- make_smooth_dens(org_densf = weighted_mixtures[[use_activity]]) # Process Model

seg_starts <- c(.1, .4, .7)
seg_len <- 0.2

ac_features <- NULL

for (use_feature in c("area", "corr", "jsd")) {
	ac_features <- rbind(ac_features, loadResultsOrCompute(file = paste0("../results/ac_features-", use_feature, ".rds"), computeExpr = {
		doWithParallelCluster(expr = {
			library(foreach)
			foreach::foreach(seed = seq(from = 1337, length.out = 1e4), .combine = rbind, .inorder = FALSE) %dopar% {
				proc = get_smoothed_curve(seed = seed, npoints = 14, span = 0.35)
				
				temp <- matrix(nrow = 1, data = sapply(X = 1:length(seg_starts), FUN = function(seg_idx) {
					# the segment
					a <- seg_starts[seg_idx]
					b <- a + seg_len
					use_x <- seq(from = a, to = b, length.out = 500)
					
					if (use_feature == "area") {
						# Area between:
						mean(abs(pm(use_x) - proc(use_x))) * (b - a)
					} else if (use_feature == "corr") {
						# Correlation:
						temp <- suppressWarnings({
							stats::cor(x = pm(use_x), y = proc(use_x))
						})
						if (is.na(temp)) 0 else temp
					} else if (use_feature == "jsd") {
						# JSD:
						jsd_segment(p = pm, q = proc, ext = c(a, b))
					}
				}))
				
				cbind(data.frame(
					seed = seed,
					feat = use_feature,
					seg_len = seg_len,
					seg_starts = paste0(seg_starts, collapse = ","),
					stringsAsFactors = FALSE
				), `colnames<-`(x = temp, value = paste0("seg_", 1:length(seg_starts))))
			}
		})
	}))
}
```


In Figure \ref{fig:pdf-cdf-approx} we show the PDF and CDF as approximated using kernel density estimation with Gaussian kernels.
The bandwidth used is "SJ" [@sj1991bandwidth].
In that figure, the number of samples increases with each row, from $5$ to $10,000$.
While for relatively few sample the PDF/CDF change more drastically, the differences between large and very large sample sizes increasingly vanish.


```{r pdf-cdf-approx, fig.cap="PDF and CDF for the three features as they get approximated with increasing sample sizes using Gaussian KDE.", fig.width=14, fig.height=18, warning=FALSE, echo=FALSE}
feat_seg <- c("area" = 3, "corr" = 1, "jsd" = 2)


par(mfrow = c(8, 6), mar = c(4,4,2,1))
set.seed(1)
for (n in c(10, 15, 20, 30, 100, 400, 1000, 1e4)) {
	for (f in names(feat_seg)) {
		temp <- ac_features[ac_features$feat == f,][[paste0("seg_", feat_seg[f])]]
		temp <- sample(x = temp, size = n, replace = FALSE)
		if (f == "jsd") {
			temp <- -log(temp)
			f <- paste0("-log(", f, ")")
		}
		
		use_bw <- bw.SJ(x = temp)
		tempdens <- density(x = temp, bw = use_bw)
		plot(tempdens, main = paste0(f, ", PDF"))
		grid()
		
		temp_cdf <- Vectorize(function(x, h = use_bw) 1 - 1 / length(temp) * sum(pnorm((x - temp) / h)))
		curve2(temp_cdf, min(tempdens$x), max(tempdens$x), main = paste0(f, ", CCDF"))
		grid()
	}
}
```

### Summary

Here, we will generate two figures.
In the first one, we show the Automatic Calibration of the three metrics after a fixed number of epochs.
In the second, we summarize Figure \ref{fig:pdf-cdf-approx} into a new figure to be used in the dissertation.


```{r}
# In this function, we create the AC plot after n observed processes.
make_ac_plot <- function(n) {
	curve2(func = pm, from = 0, to = 1, lwd = 2, col = "#FF4A36", main = paste0("Automatic Calibration after ", n, " epochs."), xlab = "Relative Project Time", ylab = "Relative Likelihood", ylim = c(.05, 2.15))
	
	polygon_colors <- RColorBrewer::brewer.pal(12, "Set3")[c(4,5,7)]
	
	for (idx in 1:n) {
		rp <- get_smoothed_curve(seed = idx, npoints = 14, span = 0.35)
		for (fidx in 1:length(feat_seg)) {
			f <- names(feat_seg)[fidx]
			a <- seg_starts[feat_seg[f]]
			b <- a + seg_len
			use_x <- c(seq(from = a, to = b, length.out = 25))
			use_x <- c(use_x, rev(use_x))
			
			polygon(x = use_x, y = c(
	      sapply(X = head(use_x, length(use_x) / 2), FUN = function(x) {
	        max(pm(x), rp(x))
	      }),
	      sapply(X = tail(use_x, length(use_x) / 2), FUN = function(x) {
	        min(pm(x), rp(x))
	      })
	    ), col = if (idx == n) polygon_colors[fidx] else "#eeeeeef8", density = if (idx == n) 30 else 100, angle = if (idx == n) 20 else 0)
		}
	}
	
	for (idx in 1:n) {
		rp <- get_smoothed_curve(seed = idx, npoints = 14, span = 0.35)
		curve2(func = rp, from = 0, to = 1, lwd = if (idx == n) 2 else 1.25, add = TRUE, col = if (idx == n) "#444444" else "#bbbbbb")
	}
  
  # Let's always over-draw the original proc:
  for (fidx in 1:length(feat_seg)) {
  	f <- names(feat_seg)[fidx]
		a <- seg_starts[feat_seg[f]]
		b <- a + seg_len
		abline(v = c(a, b), lty = 5, col = "#dddddd")
  }
  curve2(pm, 0, 1, lwd = 3, col = "#FF4A36", add = TRUE)
  grid()
}
```


In Figure \ref{fig:ac-procs} we show how the automatic calibration may look after some epochs.
In each epoch, a random process is drawn and the deviations against the process model are computed.


```{r ac-procs, fig.width=12, fig.height=7, fig.cap="The process of Automatic Calibration, where the process model \textsf{REQ} and deviations are calibrated using random processes."}
make_ac_plot(50)
```

Let's also make the summary of Figure \ref{fig:pdf-cdf-approx}.
We will group each deviation's PDF and CDF into one plot and then show three to four rows of plots after some $n$ random processes.
The result is shown in Figure \ref{fig:pdf-cdf-approx-summary}.


```{r pdf-cdf-approx-summary, echo=FALSE, fig.width=11, fig.height=7, fig.cap="Summary of the PDF and CDF for the three features for samples sizes of $n=\\{10,20,50,10^4\\}$."}
set.seed(1)
use_n <- c(10, 20, 50, 10000)
par(mfrow = c(length(use_n),3), mar = c(1,3,2,3))
xlims <- list(
	area = c(-0.01, 0.25),
	corr = c(-1.1, 1.1),
	jsd = c(2.5, 10))


for (nidx in 1:length(use_n)) {
	n <- use_n[nidx]
	for (fidx in 1:length(feat_seg)) {
		f <- names(sort(feat_seg))[fidx]
		use_xlim <- xlims[[f]]
		temp <- ac_features[ac_features$feat == f,][[paste0("seg_", feat_seg[f])]]
		temp <- sample(x = temp, size = n, replace = FALSE)
		if (f == "jsd") {
			temp <- -log(temp)
			f <- paste0("-log(", f, ")")
		}
		
		use_bw <- bw.SJ(x = temp)
		tempdens <- density(x = temp, bw = use_bw)
		
		if (nidx == length(use_n)) {
			par(mar = c(2,3,2,3))
		}
		
		plot.new()
		plot.window(xlim=use_xlim, ylim=range(tempdens$y))
		lines(x = tempdens$x, y = tempdens$y, lwd = 1.75)
		axis(1, col.axis="black")
		axis(2, col.axis="black")
		box()
		
		# CDF:
		plot.window(xlim=use_xlim, ylim=c(0,1))
		temp_kde_cdf <- Vectorize(function(x, h = tempdens$bw)  1 / length(temp) * sum(pnorm((x - temp) / h)))
		lines(x = tempdens$x, y = temp_kde_cdf(tempdens$x), col="darkred", lwd = 2)
		axis(4, col.axis="darkred")
		
		if (nidx == 1) {
			title(f)
		}
		grid()
		#mtext("1st y axis", side = 2, las=3, line=3)
		#mtext("2nd y axis", side = 4, las=3, line=3, col="darkred")
	}
}
```





## Dataset Preparation

We want to show the original dataset (using the raw features) and the same dataset, processed into scores using the CCDFs from the previous calibration.
We show only a simplified dataset, in which we measure the correlation on the first segment ($[0.1,0.3]$), the symmetric Jensen--Shannon divergence on the second segment ($[0.4,0.6]$), and the area between the curves on the third segment ($[0.7, 0.9]$).
Also, the dataset will only include the chosen activity, $\mathsf{REQ}$, and will be generated for all $15$ projects.


We will import the issue-tracking project data and create proper densities first, before creating the dataset.

```{r}
pnames_it <- paste0("Project", 1:15)
projects_it <- loadResultsOrCompute(file = "../data/projects_it.rds", computeExpr = {
	projects_it <- list()
	
	set.seed(1)
	for (pname in pnames_it) {
		temp <- readxl::read_excel(path = "../data/FD_issue-based_detection.xlsx", sheet = pname)
		use_y <- as.numeric(temp[[tolower(use_activity)]])
		use_y[is.na(use_y)] <- 0
		tempdens <- density(x = temp$`time%`, weights = use_y / sum(use_y), bw = "SJ", cut = TRUE)
		rej_samples <- rejection_sampling(x_data = tempdens$x, y_data = tempdens$y, num_x = 1e5)
		
		templ <- list()
		templ[[use_activity]] <- gaussian_kde_pdf(rej_samples)
		projects_it[[pname]] <- templ
	}
	projects_it
})
```


### Create the Dataset

Now as for the actual dataset, we will compute the same metrics on the same segments as were used for the automatic calibration.
Here, however, we compute these for every project's activity against the process model used in AC.

```{r}
use_feats <- c("area", "corr", "jsd")
ac_dataset <- loadResultsOrCompute(file = "../data/ac_dataset.csv", computeExpr = {
	ac_dataset <- data.frame(matrix(nrow = 0, ncol = 3))

	for (pname in names(projects_it)) {
		proc <- projects_it[[pname]]$REQ
		tempm <- matrix(nrow = 1, ncol = 3)
		
		for (f_idx in 1:length(use_feats)) {
			f <- use_feats[f_idx]
			# the segment
			seg_idx <- feat_seg[f]
			a <- seg_starts[seg_idx]
			b <- a + seg_len
			use_x <- seq(from = a, to = b, length.out = 500)
			
			tempm[1, f_idx] <- if (f == "area") {
				mean(abs(pm(use_x) - proc(use_x))) * (b - a)
			} else if (f == "corr") {
				temp <- suppressWarnings({
					stats::cor(x = pm(use_x), y = proc(use_x))
				})
				if (is.na(temp)) 0 else temp
			} else if (f == "jsd") {
				jsd_segment(p = pm, q = proc, ext = c(a, b))
			}
		}
		
		ac_dataset <- rbind(ac_dataset, tempm)
	}
	`colnames<-`(x = ac_dataset, value = use_feats)
})
```


Let's show the z-standardized data in Table \ref{tab:ac-dataset-zstd}.
These are the raw features for each project, with mean equal to zero and unit-variance (z-standardized).
It becomes apparent that we cannot comprehend this data at all, but it is often fed into regression models in exactly this shape.


```{r echo=FALSE}
temp <- as.data.frame(scale(ac_dataset))
rownames(temp) <- pnames_it

if (interactive()) {
  temp
} else {
  knitr::kable(
    x = temp,
    booktabs = TRUE,
    caption = paste0("Z-standardized data of three features for the activity ", use_activity, "."),
    label = "ac-dataset-zstd"
  )
}
```


### Rank-transform

We have previously carried out the automatic calibration and collected many observations for each feature in each of the defined segments.
We are now ready to extract the CCDFs and transform the original dataset into a dataset of scores.
Let's create the CCDFs using each feature's ideal value first.

```{r}
feat_ideal <- c("area" = 0, "corr" = 1, "jsd" = max(-log(ac_features[ac_features$feat == "jsd", paste0("seg_", feat_seg["jsd"])])))
ac_ccdfs <- list()

for (f in use_feats) {
	ac_ccdfs[[f]] <- (function() {
		temp <- ac_features[ac_features$feat == f,][[paste0("seg_", feat_seg[f])]]
		if (f == "jsd") {
			temp <- -log(temp)
		}
		# Transform using ideal value:
		temp <- abs(temp - feat_ideal[f])
		use_bw <- bw.SJ(x = temp)
		Vectorize(function(x, h = use_bw) 1 - (1 / length(temp) * sum(pnorm((x - temp) / h))))
	})()
}
```


```{r ccdfs, echo=FALSE, fig.width=12, fig.height=4, fig.cap="CCDFs for the three features as calibrated by the automatic calibration."}
par(mfrow = c(1,3))

for (f in use_feats) {
	x_max <- nloptr::nloptr(
    x0 = mean(1e-3),
    opts = list(
      maxeval = 2e2,
      algorithm = "NLOPT_GN_DIRECT_L_RAND"),
    eval_f = function(x) (ac_ccdfs[[f]](x) - 1e-3)^2,
    lb = 0,
    ub = 10
  )$solution
	
	curve2(func = ac_ccdfs[[f]], from = 0, to = x_max, lw = 1.25, main = paste0("CCDF of ", if (f != "jsd") f else paste0("-log(", f, ")")))
	grid()
	curve2(func = ac_ccdfs[[f]], from = 0, to = x_max, lw = 1.25, add = TRUE)
}
```


Figure \ref{fig:ccdfs} shows the CCDFs associated with each feature.
They are used to transform the dataset into scores, where a higher score indicates a lower distance to the ideal value.
This is shown in Table \ref{tab:ac-dataset-scores}.
This table allows us now for each project and each metric, to assess how good it matches the process model in the metric's segment.


```{r}
ac_dataset_scores <- `rownames<-`(x = data.frame(
	area = ac_ccdfs$area(abs(feat_ideal["area"] - ac_dataset$area)),
	corr = ac_ccdfs$corr(abs(feat_ideal["corr"] - ac_dataset$corr)),
	jsd = ac_ccdfs$jsd(abs(feat_ideal["jsd"] - -log(ac_dataset$jsd)))), value = pnames_it)
```

```{r echo=FALSE}
if (interactive()) {
  ac_dataset_scores
} else {
  knitr::kable(
    x = ac_dataset_scores,
    booktabs = TRUE,
    caption = "The previous dataset where all features have been transformed to scores.",
    label = "ac-dataset-scores"
  )
}
```


## Regression Tests and Rank-transforms

Let's test whether a simple linear model works better with the original data or with scores.
My hypothesis is that it will work better (that means, lower average validation error) with scores.
We will test this using Leave-one-out Cross-validation [@lachenbruch1968Loocv].

First, we have to load the ground truth.

```{r}
ground_truth_all <- rbind(
  read.csv(file = "../data/ground-truth.csv", sep = ";"),
  read.csv(file = "../data/ground-truth_2nd-batch.csv", sep = ";"))
```



### LOOCV Using the Ordinary Dataset

Here we train a model using the original dataset.
We will pre-process the raw features simply by standardizing them.
Table \ref{tab:varimp-averaged} shows the averaged variable importance.

```{r}
predicted <- c()
dataset <- cbind(ac_dataset, data.frame(gt = ground_truth_all$consensus / 10))
var_imp <- NULL

for (pId in 1:15) {
	train <- dataset[setdiff(1:15, pId),] # LOO dataset
	valid <- dataset[pId,]
	
	pre_proc <- caret::preProcess(train[, "gt" != colnames(train)], method = c("center", "scale"))
	train <- predict(pre_proc, train)
	valid <- predict(pre_proc, valid)
	
	model <- lm(formula = gt~., data = train)
	#model <- randomForest::randomForest(gt~., train)
	#model <- neuralnet::neuralnet(formula = gt~., data = train, hidden = c(2,2))
	predicted <- c(predicted, predict(model, valid))
	
	temp <- caret::varImp(model) / sum(caret::varImp(model))
	# We'll average the variable importance later
	var_imp <- if (is.null(var_imp)) temp else var_imp + temp
}

print(model$coefficients)
print(sqrt(mean((predicted - dataset$gt)^2)))
var_imp <- var_imp / sum(var_imp)
```

```{r echo=FALSE}
temp <- data.frame(var_imp)

if (interactive()) {
  temp
} else {
  knitr::kable(
    x = temp,
    booktabs = TRUE,
    caption = "Averaged variable importance for the trained model.",
    label = "varimp-averaged"
  )
}
```


### LOOCV Using the Scores Dataset

Let's do the leave-one-out training for every project.
Then, in Table \ref{tab:varimp-averaged-scores}, we show the variable importance of the trained model.

```{r}
predicted <- c()
dataset <- cbind(ac_dataset_scores, data.frame(gt = ground_truth_all$consensus / 10))
var_imp <- NULL

for (pId in 1:15) {
	train <- dataset[setdiff(1:15, pId),] # LOO dataset
	valid <- dataset[pId,]
	
	# This has no effect on a linear model.
	#pre_proc <- caret::preProcess(train[, "gt" != colnames(train)], method = c("center", "scale"))
	#train <- predict(pre_proc, train)
	#valid <- predict(pre_proc, valid)
	
	model <- lm(formula = gt~., data = train)
	#model <- randomForest::randomForest(gt~., train)
	#model <- neuralnet::neuralnet(formula = gt~., data = train, hidden = c(2,2))
	predicted <- c(predicted, predict(model, valid))
	
	temp <- caret::varImp(model) / sum(caret::varImp(model))
	# We'll average the variable importance later
	var_imp <- if (is.null(var_imp)) temp else var_imp + temp
}

print(model$coefficients)
print(sqrt(mean((predicted - dataset$gt)^2)))
var_imp <- var_imp / sum(var_imp)
```

```{r echo=FALSE}
temp <- data.frame(var_imp)

if (interactive()) {
  temp
} else {
  knitr::kable(
    x = temp,
    booktabs = TRUE,
    caption = "Averaged variable importance for the model trained using scores instead of standardized features.",
    label = "varimp-averaged-scores"
  )
}
```



### Rank-transforms

Let's demonstrate raw features vs. z-standardized features vs. scores for one example.
We will pick one project and learn the scaling on the others.
Then, we'll train a predictive model and evaluate the result using scores and variable importance.

```{r warning=FALSE}
dataset <- cbind(ac_dataset, data.frame(gt = ground_truth_all$consensus / 10))
use_pid <- 15 # 15, 7, 12
train <- dataset[setdiff(1:15, use_pid),]
valid <- dataset[use_pid,]

set.seed(1)
model <- randomForest::randomForest(gt~., train, importance=TRUE, localImp=TRUE, mtry=3)
c(predict(model, valid), valid$gt)
```

And the importances are:

```{r}
tempdf <- as.data.frame(model$importance)
(temp_imp <- `names<-`(x = abs(tempdf$`%IncMSE`) / sum(abs(tempdf$`%IncMSE`)), values = rownames(tempdf)))
```

```{r pm-vs-p, echo=FALSE, fig.width=10, fig.height=5, fig.cap="Correlation, Jensen--Shannon divergence, and area between for the process model (REQ) vs. a process."}
plot.new()
plot.window(xlim=c(0,1), ylim=c(0.2,1.8))
axis(1)
axis(2)
mtext("Relative Project Time", side = 1, las=1, line=3)
mtext("Relative Likelihood", side = 2, las=3, line=3)
box()
grid()

tempf <- projects_it[[paste0("Project", use_pid)]]
tempf_upper <- Vectorize(function(x) max(pm(x), tempf$REQ(x)))
tempf_lower <- Vectorize(function(x) min(pm(x), tempf$REQ(x)))

use_cols <- RColorBrewer::brewer.pal(12, "Set3")[c(4,5,7)]

for (f in names(feat_seg)) {
	a <- seg_starts[feat_seg[f]]
	b <- a + seg_len
	use_x <- seq(from = a, to = b, length.out = 20)
	use_y <- c(tempf_upper(use_x), tempf_lower(rev(use_x)))
	use_x <- c(use_x, rev(use_x))
	rect(xleft = a, ytop = 2, xright = b, ybottom = 0, col = "#00000003")
	polygon(x = use_x, y = use_y, density = 30, angle = 20, col = use_cols[feat_seg[f]])
}

curve2(func = pm, 0, 1, col = "red", lwd = 1.9, add = TRUE)
curve2(func = projects_it[[paste0("Project", use_pid)]]$REQ, 0, 1, lwd = 1.5, add = TRUE)
abline(v = c(.1,.3,.4,.6,.7,.9), lty = 5, col = "#dddddd")


legend(x = 0.55, y = 1.8, legend = c("Correlation", "Jensen--Shannon Divergence", "Area between", "PM"),
			 angle = c(rep(20, 3), 0), density = c(rep(30, 3), 0), lty = c(rep(NA, 3), 1), lwd = c(rep(NA, 3), 2), border = c(rep("#000000", 3), NA), fill = c(use_cols[1:3], NA), bg = "white", x.intersp = c(1,1,1,1), col = c(rep(NA, 3), "red"))
```

Figure \ref{fig:pm-vs-p} shows an example of measuring the deviation between a process model and a process.
Table \ref{tab:ac-std-vs-scores} then shows the original feature's data, the scaled data, and the scores for this example.

```{r echo=FALSE}
train_scaler <- attributes(scale(x = train[, 1:3]))

temp <- `rownames<-`(x = rbind(
	valid[, 1:3],
	`colnames<-`(x = matrix(as.numeric(valid[1, 1:3] - train_scaler$`scaled:center`), nrow = 1) / matrix(train_scaler$`scaled:scale`, nrow = 1), value = names(valid[, 1:3])),
	ac_dataset_scores[use_pid,]
), value = c("Raw Feature", "Z-scaled", "As Score"))


if (interactive()) {
  temp
} else {
  knitr::kable(
    x = temp,
    booktabs = TRUE,
    caption = "Raw features, Z-standardized features, and Scores for the same deviations.",
    label = "ac-std-vs-scores"
  )
}
```


```{r}
# We get the "mean" score using the sum, because the weights already sum to 1!
sum(`names<-`(x = as.numeric(temp_imp * ac_dataset_scores[use_pid,]), value = colnames(ac_dataset_scores)))
```

```{r}
# These are the weighted scores:
`names<-`(x = as.numeric(temp_imp * ac_dataset_scores[use_pid,]), value = colnames(ac_dataset_scores))
```




\clearpage

# References {-}

<div id="refs"></div>












